---
title: "Homework 2: Data Wrangling"
output: github_document
---

## Problem 1

This is the start of Homework 2 for P8105: Data Science 1. First, we are going to attach the packages needed for this homework.  

```{r Loading Packages}
library(tidyverse)
library(readxl)
```

Next, we are going to import and clean the NYC Transit Subway dataset.  

```{r Importing Data}
transit_df = 
  read_csv(file = "NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |>
janitor::clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude, starts_with("route"),
    entry, exit_only, vending, entrance_type, ada) |>
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```


How many distinct stations are there? 

```{r}
transit_df |> 
  select(station_name, line) |>
  distinct()
```

How many stations are ADA complaint? 84 stations. 
```{r}
transit_df |> 
  filter(ada == "TRUE") |>
  select(station_name, line) |> 
  distinct()
```


What proportion of station entrances/exits without vending allow entrance? 

```{r}
transit_df |> 
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```

How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA complaint? 

60 stations serve the A train. Of the stations that serve the A train, none are ADA complaint. 

```{r}
transit_df |> 
  pivot_longer(
    route1:route11, 
    names_to = "route_num", 
    values_to = "route") |>
  filter(route == "A") |>
  select(station_name, line) |>
  distinct()

transit_df |> 
  pivot_longer(
    route1:route11,
    names_to = "route") |>
  filter(route == "A", ada == TRUE) |>
  select(station_name, line) |>
  distinct()
```

## Problem 2

This is the start of Problem 2. First, we are going to import the Mr. Trash Wheel dataset.We will clean the variable names, and convert the sports_balls variable into an integer variable. 

```{r}
library(readxl)

trash_df = 
  read_excel("202409_Trash_Wheel_Collection_Data.xlsx",
    na = c("NA", "")) |> 
  janitor::clean_names() |> 
  mutate(sports_balls = as.integer(round(sports_balls))) |>
  mutate(year = as.numeric(year)) |>
  mutate(mr_trash_wheel = plastic_bottles + plastic_bags) 
```

Next, we are importing the Professor Trash Wheel sheet from the Trash Wheel dataset.

```{r}
professor_df = 
  read_excel("202409_Trash_Wheel_Collection_Data.xlsx", sheet = "Professor Trash Wheel",
    na = c("NA", "")) |>
  janitor::clean_names() |>
  mutate(professor_wheel = plastic_bottles + plastic_bags)
```

Now, we are importing the Gwynnda Trash Wheel sheet from the Trash Wheel dataset.

```{r}
Gwynnda_df = 
  read_excel("202409_Trash_Wheel_Collection_Data.xlsx", sheet = "Gwynnda Trash Wheel",
    na = c("NA", "")) |>
  janitor::clean_names() |> 
mutate(Gwynnda_wheel = plastic_bottles + plastic_bags)
```

As the next step to this problem, we are going to combine the Mr. Trash Wheel, Professor Trash Wheel, and the Gwynnda Trash Wheel datasets. 

```{r Binding}
trash_tidy = 
  bind_rows(trash_df, professor_df, Gwynnda_df) |> 
  janitor:: clean_names() |> 
  pivot_longer(
    mr_trash_wheel:gwynnda_wheel,
    names_to = "Dataset_type", 
    values_to = "plastic")
```

Our combined data has 3,114 observations and 18 variables.Some key variables are weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, and glass_bottles. 

## Problem 3 

First, I am going to import and clean the individual bakers dataset. 

```{r}
bakers_df = 
  read_csv("gbb_datasets/bakers.csv",
    na = c("NA", "", ".")) |>
  janitor::clean_names()
```

Next, we are importing and cleaning the bakes dataset. 
```{r}
bakes_df = 
  read_csv("gbb_datasets/bakes.csv",
    na = c("NA", "", ".")) |>
  janitor::clean_names()
```

Finally, we are importing and cleaning the results dataset. For the results dataset, I renamed the columns "series", "episode", "baker", "technical", and "results." Next, I am going to remove the first three rows of the results dataset, and then change the series and episode variables to numeric. Should I have changed the results variable to something other than character? 
```{r}
results_df = 
  read_csv("gbb_datasets/results.csv", 
    na = c("NA", "", "."), col_names = c("series", "episode", "baker", "technical","result" ), col_types = cols('series' = col_double(), 'episode' = col_double(), 'technical' = col_double())) |>
  janitor::clean_names() 
results_df <- results_df[-c(1,2,3),]
```

Now, I am going to merge the datasets together. 

```{r}
bake_tidy = 
  anti_join(bakers_df, bakes_df, by= "series")
```

